---
title: "ADV_R_project"
author: "Evrim Bilgen (434281) & Gamze Kopal (436823)"
date: "1/15/2022"
output: html_document
---



## References
● ILO
● OECD.Stat
● The World Bank
● Heckman J. and Killingsworth M. (1986) ‘Female Labor Supply: A Survey. in
Handbook of Labor Economics
● Princeton University
● Kaggle


## R Markdown

```{r}

library(readr)
library(dplyr)
library(ggplot2)
library(readr)
library(leaflet)
library(gapminder)
library(stringr)
library(lubridate)
library(tidyverse)
library(cowplot)
library(colorspace)
library(ggrepel)
library(viridis)
library(gganimate)
library(ggplot2)
library(dplyr)
library(maps)
library(viridis)
library(directlabels)
library(ggrepel)
library(maps)
library(RColorBrewer)
library(forcats)
library(ggrepel)
library(gridExtra)

hrbrthemes::import_roboto_condensed()

# Upload data
lfp <- read.csv('/Users/ewrimmm/Desktop/adR/Data/recent-ILO-LFP.csv')
lfp_1980 <- read.csv('/Users/ewrimmm/Desktop/adR/Data/female-labor-force-participation-rates-latest-vs-1980.csv')
ratio <- read.csv('/Users/ewrimmm/Desktop/adR/Data/ratio-of-female-to-male-labor-force-participation-rates-ILO-WDI.csv')
informal <- read.csv('/Users/ewrimmm/Desktop/adR/Data/women-in-informal-employment-as-share-of-female-employment.csv')
unpaid <- read.csv('/Users/ewrimmm/Desktop/adR/Data/female-to-male-ratio-of-time-devoted-to-unpaid-care-work.csv')
scarce <- read.csv('/Users/ewrimmm/Desktop/adR/Data/share-who-agree-with-the-statement-when-jobs-are-scarce-men-should-have-more-right-to-a-job-than-women.csv')
fertility <- read.csv('/Users/ewrimmm/Desktop/adR/Data/fertility-and-female-labor-force-participation.csv')

crime <- read.csv('/Users/ewrimmm/Desktop/adR/Data/crime_female.csv')

manager <- read.csv('/Users/ewrimmm/Desktop/adR/Data/female_managers.csv')

emp_unem <- read.csv('/Users/ewrimmm/Desktop/adR/Data/emp&uneteR.csv')

sector <- read.csv('/Users/ewrimmm/Desktop/adR/Data/bysex&sectors.csv')


```


# Female Labor Force Participation Rates - Regions
```{r}

regions <- c("Sub-Saharan Africa",
             "East Asia & Pacific",
             "North America",
             "Latin America & Caribbean",
             "European Union",
             "Central Europe and the Baltics",
             "South Asia",
             "Middle East & North Africa")

regions_lfp <- subset(lfp, Entity %in% regions)

regions_lfp$label <- NA
regions_lfp$label[which(regions_lfp$Year == max(regions_lfp$Year))] <-
  regions_lfp$Entity[which(regions_lfp$Year == max(regions_lfp$Year))]


ggplot(data=regions_lfp, aes(x=Year, y=LFP, group=Entity, color=Entity)) +
  geom_line(linetype = "dashed")+
  geom_line(aes(color=Entity))+
  geom_point(aes(color=Entity))+
  ggtitle("Female Labor Force Participation Rates, 1990-2019") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))+
  geom_label_repel(aes(label = label),
                   nudge_x = 1,
                   na.rm = TRUE) +
  theme(legend.position = "none")
```

# Female Labor Force Participation Rates 1980 vs 2020
```{r}
ggplot(data=lfp_1980, aes(x=LFP_2020, y=LFP_1980, color=Entity, label=Entity))+
  geom_dotplot(binaxis='y', stackdir='center', method = 'dotdensity', aes(fill=Entity))+
  geom_text(hjust=0, vjust=0)+
  ggtitle("Female Labor Force Participation Rates, 1980 vs 2020") +
  xlab("Female labor force participation rate (%) in 2020") + ylab("Female labor force participation rate (%) in 1980")+
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position = "none")
```

# Female to Male Ratio in World Map
```{r}
world_map <- map_data(map = "world")
world_map$region <- iso.alpha(world_map$region, n=3)

ratio$count_group <- cut(ratio$FTM,
                      breaks = c(-Inf, 30, 40, 50, 60, 70, 80, 90, 100, Inf),
                      labels = c("30%", "40%", "50%", "60%","70%","80%", "90%", "100%", "110%"))


ggplot(ratio) +
  geom_map(aes(map_id = Code, fill = fct_rev(count_group)), map = world_map) +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group), colour = 'black', fill = NA) +
  expand_limits(x = world_map$long, y = world_map$lat) +
  scale_fill_manual(name = "Ratios", values = rev(brewer.pal(9, name = "Purples"))) +
  theme_void() +
  coord_fixed() +
  ggtitle("The Female-to-Male Ratio of Labor Force Participation Rates %, 2019") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position = "bottom")
```



# Informal Employment
```{r}
informal$informal= round(informal$informal,1)

ggplot(data=informal, aes(x=informal, y= reorder(Entity, informal))) +
  geom_bar(stat="identity", aes(fill=Entity)) +
  geom_text(aes(label=informal), position=position_dodge(width=0.9), hjust=-0.1)+
  xlab("Informal Employment Share") + ylab("Country")+
  ggtitle("Informal employment as share of female employment, 2017") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position = "none")
```



# Unpaid Care Work
```{r}
colnames(unpaid)[4] <- "unpaid"
unpaid <- unpaid[with(unpaid,order(-unpaid)),]

unpaid <- unpaid[3:12,]

ggplot(data=unpaid, aes(x=unpaid, y= reorder(Entity, -unpaid))) +
  geom_bar(stat="identity", aes(fill=Entity)) +
  geom_text(aes(label=unpaid), position=position_dodge(width=0.9), hjust=-0.1)+
  xlab("Female-to-male unpaid care work ratio") + ylab("Country")+
  ggtitle("Female-to-male ratio of time devoted to unpaid care work, 2014") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position = "none")
```



# Fertility
```{r}
ggplot(data=fertility, aes(x=birth, y=LFP, color=Entity, label=Entity))+
  geom_dotplot(binaxis='y', stackdir='center', method = 'dotdensity', aes(fill=Entity), binwidth = 1)+
  geom_text(hjust=0, vjust=0)+
  ggtitle("Labor Force Participation Rates and Births, 1980 vs 2020") +
  ylab("Female labor force participation rate (%) in 2020") + xlab("Live birth per woman")+
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position = "none")
```



# Men should have more right in employment(?)
```{r}
colnames(scarce)[4] <- "Yes"

scarce <- scarce[with(scarce,order(-Yes)),]

#scarce <- scarce[1:20,]

ggplot(data = scarce, aes(x = reorder(Entity, Yes), y = Yes)) +
  geom_bar(stat = 'identity', fill = 'purple', color = 'violet') +
  ggtitle('Share agrees men should have more right in employment') +
  xlab("Country") + ylab("Yes %")+
  theme_minimal() +
  geom_text(aes(label = Yes), vjust = -0.5, color = 'black', size = 3.2)+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```

#reshaping Time column
```{r}
emp_unem$Time <- str_replace_all(emp_unem$Time,"Q1-","")
emp_unem$Time <- str_replace_all(emp_unem$Time,"Q2-","")
emp_unem$Time <- str_replace_all(emp_unem$Time,"Q3-","")
emp_unem$Time <- str_replace_all(emp_unem$Time,"Q4-","")
```


#reshaping IND column
```{r}
emp_unem$IND[emp_unem$IND == "EMP4_U"] <- "U" 
emp_unem$IND[emp_unem$IND == "EMP4_E"] <- "E" 

```


#rename Age Group column
```{r}
names(emp_unem)[names(emp_unem) == 'Age.Group'] <- 'Age_Group'
```



#Disturbution of Age Group
```{r}
rs3 <- emp_unem %>%
  group_by(Age_Group) %>%
  summarise(number = n())

ggplot(data = rs3, aes(x = Age_Group, y = number)) +
  geom_bar(stat = 'identity', fill = 'purple', color = 'violet') +
  ggtitle('Disturbution of Age Group') +
  xlab('Age Group') +
  ylab('Number') +
  theme_minimal() +
  geom_text(aes(label = number), vjust = -0.5, color = 'black', size = 2)
```



#Age Group Disturbution by Gender
```{r}
rs4 <- emp_unem %>%
  group_by(Age_Group, Sex) %>%
  summarise(number = length(Age_Group))

g1 <- ggplot(data = rs4, aes(x = Age_Group, y = number, fill = factor(Sex))) +
  geom_bar(na.rm = TRUE, position = "dodge", width = 0.5, stat = "identity") +
  ggtitle('Age Group Disturbution by Gender', subtitle='Employed&Unemployed') +
  xlab('Age') +
  ylab('Population') +
  labs(fill = 'Gender') +
  theme_minimal()

```

#Finding unemployed by gender
```{r}
emp_unem_2 <- emp_unem %>%
  filter(IND=='U',Time>2010)%>%
  group_by(Age_Group,Sex,COU) %>%
  summarise(number = n())

g2 <- ggplot(data = emp_unem_2, aes(x = Age_Group, y = number, fill = factor(Sex))) +
  geom_bar(na.rm = TRUE, position = "dodge", width = 0.5, stat = "identity") +
  ggtitle('Age Group Disturbution by Gender', subtitle='Unemployed (Time range > 2010)') +
  xlab('Age') +
  ylab('Population') +
  labs(fill = 'Gender') +
  theme_minimal()

```


#Finding unemployed women by countries
```{r}

BigCountries <- c('TUR','FRA','POL','ESP','ISR','RUS','JPN','NOR','DEU','NLD','GBR')



g3 <- emp_unem %>%
  filter(IND=='U',Time>2010, COU %in% BigCountries) %>%
  ggplot(aes(x=COU)) +
  geom_bar(fill='dark blue')  +
  # add theme_minimal()
  theme_minimal() +
  coord_flip()
```

```{r}
grid.arrange(g1, g2, g3)
```

#Sector csv
```{r}
ds <- sector %>%
  filter(Time > 2011)


WesternCountries <- c('CAN','DNK','CZE','DEU','FRA','GRC','HUN','ITA','NOR','PRT')

ds <- sector %>%
  filter(COU %in% WesternCountries,Time > 2018, Sex=='Women')

#country_cols <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#999999")
```


```{r}
ggplot(ds, aes(x=COU, y=Value)) +
  geom_segment( aes(x=COU, xend=COU, y=0, yend=Value), color="grey") +
  geom_point( color="purple", size=4) +
  theme_light() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.border = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  ggtitle('Participate rate in the industries last 10 years') +
  xlab("Western Countries") +
  ylab("Participate rate")
```



```{r}

sector_2 <- sector %>%
  filter(Time>2007)%>%
  group_by(Sex,Time) %>%
  summarise(number = n())


ggplot(sector_2, aes(x=Time, y=number, fill=Sex)) +
  geom_bar(stat='identity')
```


#interactive -1
```{r}
ggplot(sector_2, aes(x=Time, y=number, fill=Sex)) +
  geom_bar(stat='identity',colour="black") +
  ggtitle("Female&Male Employees in the previous 10 years") +
  ylab("Total employer") +
  theme_bw() +
  # gganimate specific bits:
  transition_states(
    Time,
    transition_length = 2,
    state_length = 1
  ) +
  ease_aes('sine-in-out')


  anim_save("sector-barplot-transition.gif")

```


#female managers

#interactive -2
```{r}
MapWorld <- map_data("world")


dd2 <- manager %>% filter(COU %in% c('DNK','CZE','DEU','FRA','GRC','HUN','ITA','NOR','POL')) 


ggplot(dd2 , aes(x=Time, y=Value, group=Country, color=COU)) +
  geom_line() +
  geom_point() +
  scale_color_viridis(discrete = TRUE) +
  transition_reveal(Time)

anim_save("Female_boss.gif")

```


#Crime
```{r}
crime_variables <- unique(crime$Variables)


crime %>% filter(Region=='Europe and Central Asia')%>%
  ggplot(aes(x = Country, y = Value,
             fill = Variables)) +
  geom_col(show.legend = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
  theme(legend.position = c(0.8, 0.2)) +
  facet_wrap(~Variables)

```
